---
title: "Linear regression - Power plant data set"
author: "FdR"
date: "29 December 2014"
output: html_document
---

This post is an attempt to use this article from [Data Perspective](http://www.dataperspective.info/2014/10/regression-analysis-using-r.html) into a R Markdown document using, Github, ggplot2, dplyr, etc. 

The article use the [data](https://archive.ics.uci.edu/ml/datasets/Combined+Cycle+Power+Plant) from a combine cycle power plant.  Once downloaded and put in the right folder, we can start working on it.
Because, it is a .xlsx, let's first install the library, then load it.

```{r}
library(xlsx)
```

We then load the file.  Because this .xlsx file has several index (tab), we have to specify which tab, we want to use.
````{r}
data = read.xlsx("Folds5x2_pp.xlsx", sheetIndex=1)
````

Let's have a quick look at our data
````{r}
View(data)
````

which is quite similar to
````{r}
head(data)
````

We can focus on the first column (AT ~ Average Temperature)
Any missing data?
````{r}
sum(is.na(data$AT))
````
The range of data?
````{r}
range(data$AT)
````

Plot the distribution of values for AT.
````{r}
library(ggplot2)
p1 = ggplot(data, aes(AT)) + geom_density(color="blue") + scale_x_continuous(breaks = seq(1,40,2))
````


You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
